---
title: "Nov 26  in class noes"
author: "Lucy Zhang"
date: "November 26, 2018"
output: 
  html_document:
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Stochastic simulations

* The outcome is determined when the math is the same, the model is static, adding sources of variation at various places in model and allowing to \emph{propagate} and influence your models

##Randomness

* There is no true randomness, only \emph{pseudo-random} generation

```{r}
set.seed(720)
x <- rnorm(1000, mean = 5, sd = 1)
mean(x)
set.seed(720)
x <- rnorm(1000, mean = 5, sd = 1)
mean(x)
set.seed(720)
x <- rnorm(1000, mean = 5, sd = 1)
mean(x)
```

* It's the same number every time if you set the seed the first time, but if you don't it will continue getting random numbers

* If you don't set the seed, the computer uses the clock on your computer

* Setting the seed is important because you want to be able to make reproducible data for other people

##Rolling the dice

```{r}
sample(1:6, size = 1 , replace = T , prob = NULL)
```

* `sample()` is a function that samples permutations
  + `x` is the thing you want to sample
  + `size` is the number of times you want to sample
  + `replace` toggles sampling with replacement
  + `prob` is the probably of each samples
  
* Usually we are interested in sampling with replacment

```{r}
replicate(100, sample(1:6, size = 1 , replace = T , prob = NULL))
```

```{r}
more_rolls <- sample(1:6, size = 1000 , replace = T , prob = NULL)
barplot(table(more_rolls))
```

* If replace = F, then you can't take a sample larger than the population (6)

##How is sampling useful

* empirical sampling methods

* Randomzation/permutation tests, shuffling the order relative to a treatment variable

* Non-parametric bootstrap -> sampling with replacement

##Modelling genetic drift

* A stochastic sampling of which alleles are transmitted from one generation to the next -> *by random chance*

```{r}
sampling <- sample(c("A", "a"), size = 40, rep = T, prob = c(0.5, 0.5))
table(sampling)
```

* Allele frequencies

```{r}
freq1 <- table(sampling)/length(sampling)
sampling <- sample(c("A", "a"), size = 40, rep = T, prob = freq1)
table(sampling)
```

##Numerical stochastic simulations

* Simulate numbers to form a probability distribution

* `runif` R uniform, generates a random number between your set ranges 

```{r}
runif(n = 1, min = 0, max = 1)
```

```{r}
ru <- runif(n = 10000, min = 0, max = 1)
head(ru)
tail(ru)
hist(ru, freq = F)
```

```{r}
random.normal.100 <- rnorm(n = 100, mean = 5, sd= 2)
par(mfrow = c(2, 2))
plot(random.normal.100)
hist(random.normal.100)
qqnorm(y = random.normal.100)
```

* `qqplot` comparing data to theoretical distribution

```{r}
random.normal.100 <- rnorm(n = 100, mean = 5, sd= 2)
thing <- replicate(100, random.normal.100)
hist(thing, freq = F)
```

##Monte Carlo simulation

* Deterministic part
```{r}
par(mfrow=c(1,1))
a = 5
b = 0.7
x <- seq(2, 20)
yfix <- a + b*x
plot(yfix ~ x, main = "Deterministic Component")
abline(a = 5, b = 0.7)
```
* Adding stochastic and random compoenents to the data

* Red = true fit 

* Blue = regression based on simulation

```{r}
y.sim.1 <- rnorm(length(x), mean = yfix, sd = 2.5)
plot(y.sim.1 ~ x, pch = 20)
abline(a = 5, b = 0.7, col = "red")
y.sim.1.lm <- lm(y.sim.1 ~ x)
abline(reg = y.sim.1.lm, lty = 2, col = "blue")

```

* Repeat the simulation 25 times

```{r}
plot(yfix ~ x, main = "Deterministic Component")
for (i in 1:25) {
  x <- seq(2, 20)
  y.sim <- rnorm(length(x), mean = yfix, sd = 2)
  points(y.sim ~ jitter(x), pch = 20, col = rgb(0, 0, 0, alpha = 0.5))
  y.sim.lm <- lm(y.sim ~ x)
  abline(reg = y.sim.lm, lty = 2, col = rgb(0, 0, 1, alpha = 0.5))
}
abline(a = 5, b = 0.7, col = "red")
```

* Generate 100000 bp sequence

```{r}
seq1 <- sample(c("A", "T", "C", "G"), 100000, replace = T, prob = rep(0.25, 4))
table(seq1)/length(seq1)
```

* Combine into a string vector

```{r}
seq1_col <- paste(seq1, collapse = "")
```

##Global regular expressions parser

```{r}
x <- gregexpr("AACTTTT", seq1_col, fixed = T, useBytes = T)
length(unlist(x))
```

* Parameters to speed up search
  + `useBytes` byte by byte comparison, works on any ASCII characters
  + `fixed` look through not everything
  
* In *Drosophila*, the genes have GC bias

* In `prob`, the smaller number is smaller probability obv!

```{r}
seq2 <- sample(c("A", "T", "C", "G"), 100000, replace = T, prob = c(0.3, 0.3, 0.2, 0.2))
table(seq2)/length(seq2)
seq2_col <- paste(seq2, collapse = "")
y <- gregexpr("AACTTTT", seq2_col, fixed = T, useBytes = T)
length(unlist(y))
```