---
title: "In Class, Nov 5, 2018"
author: "Lucy Zhang"
date: "November 5, 2018"
output: 
  html_document:
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Download dataset
```{r}
dll_data = read.csv("http://datadryad.org/bitstream/handle/10255/dryad.8377/dll.csv", header=T)
```
View your data
```{r}
summary(dll_data)
str(dll_data)
dim(dll_data)
head(dll_data)
```
Checking the data 
```{r}
head(is.na(dll_data))
sum(is.na(dll_data))
anyNA(dll_data)
```
Cleaning the data 
```{r}
dll_data_complete <- na.omit(dll_data)
dim(dll_data)
dim(dll_data_complete)
dll_data <- na.omit(dll_data)
mean(dll_data$femur)
```
Check for duplicated data using `duplicated()`
```{r}
head(duplicated(dll_data))
tail(duplicated(dll_data))
sum(duplicated(dll_data))
any(duplicated(dll_data))
dll_data[anyDuplicated(dll_data),]
```
Sampling rows and randomly picking them
```{r}
new_rows <- dll_data[sample(nrow(dll_data), size = 5, replace = T ),]
dll_data2 <- rbind(dll_data, new_rows)
str(dll_data2)
```
Then checking it again
```{r}
any(duplicated(dll_data2))
```
Find out which rows are duplicated
```{r}
dll_data2[duplicated(dll_data2),]
```
Remove data that is duplicated 
```{r}
dll_data_unique <- unique(dll_data2)
dim(dll_data_unique)
dim(dll_data2)
dim(dll_data)
```
Remove environmental variables
```{r}
rm(dll_data_complete, dll_data_unique, dll_data2)
```
Subseting the data for only rows that have `wt`
```{r}
dll_data_wt <- dll_data[dll_data$genotype == "wt",]
```
Checking the subset
```{r}
with(dll_data, table(genotype))
nrow(dll_data_wt)
levels(dll_data_wt$genotype)
```
Even though the data has no more `Dll`, it still keeps the levels. To remove the un-needed levels
```{r}
dll_data_wt <- droplevels(dll_data_wt)
levels(dll_data_wt$genotype)
```
Make a data frame for the Dll factor
```{r}
dll_data_Dll <- subset(dll_data, genotype == "Dll" )
dim(dll_data_Dll)
dll_data_Dll2 <- subset(dll_data, genotype == "Dll" , drop= T)
dim(dll_data_Dll2)
with(dll_data, table(genotype))
levels(dll_data_Dll$genotype)
dll_data_Dll <- droplevels(dll_data_Dll)
```
Make a new subset of data
```{r}
dll_data_Dll <- subset(dll_data, 
                       genotype == "Dll",
                       select = c(line, genotype, temp, SCT))
dim(dll_data_Dll)
```
Find things in your data
```{r}
matched_set <- dll_data$line %in% c("line-Sam", "line-1")
sum(matched_set)
dll_data_new_subset <- dll_data[matched_set,]
dim(dll_data_new_subset)
```
```{r}
rm(dll_data_Dll, dll_data_wt, dll_data_new_subset, matched_set)
levels(dll_data$line)
```
Factors are stored as integers in R to make it more efficient, have to convert the factor into `chr` for string manipulation
```{r}
line_str <- as.character(dll_data$line)
str(line_str)
head(line_str)

line_names <- substr(line_str, 
                     start = 6, stop= 1000000L )

head(line_names)
tail(line_names)
```
`gsub()` replaces elements in the objects specified
```{r}
str(line_str)
line_names3 <- gsub(pattern = "line-",
                    replacement = "",
                    x = line_str)
```
Makes things to lower case, `toupper()` makes things into upper case
```{r}
line_names3 <- tolower(line_names3)
head(line_names3)
tail(line_names3, n = 25)
```
Only changes the first one 
```{r}
line_names3 <- sub("sam", "SAMARKAND", line_names3)
rm(line_names2, line_names3, line_names, line_names2_mat, line_str, new_rows)
```
Extracting temperatures 
```{r}
str(dll_data$temp)
```
Generate a factor of temperature, changes the original data, making a new data set with `temp` as a factor
```{r}
temp_as_factor <- with(dll_data,
                      factor(temp, 
                             labels = c("LowTemp", "HighTemp")))
str(temp_as_factor)
```
Using `ifelse()` to generate a factor
```{r}
temp_as_factor2 <- with(dll_data, ifelse(temp == 25, "LowTemp", 
                                         ifelse(temp == 30, "HighTemp", NA)))
temp_as_factor2 <- factor(temp_as_factor2)
str(temp_as_factor2)
```
`sort` only sorts a single variable and is not designed to handle data frames. `order` should be used, highest to lowest
```{r}
head(order(dll_data$SCT))
tail(order(dll_data$SCT))
dll_data_sorted <- dll_data[order(dll_data$SCT),]
head(dll_data_sorted)
```
Change from lowest to highest -> reverse
```{r}
dll_data_sorted <- dll_data[order(-dll_data$SCT),]
head(dll_data_sorted)
dll_data_sorted <- dll_data[order(dll_data$SCT, decreasing = T),]
dll_data_sorted <- dll_data[order(dll_data$tibia),]
head(dll_data_sorted)
```
Secondary, tertiary sort
```{r}
dll_data_sorted <- dll_data[order(dll_data$SCT, dll_data$temp),]
head(dll_data_sorted)
```
`merge` will compare two things scruitinously