---
title: "String Manipulation"
author: "Lucy Zhang"
date: "December 3, 2018"
output: 
  html_document:
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(stringr)
library(stringi)
library(tidytext)
```

**How might you get `writeLines` to print each string on a newline?**

```{r}
fly_seq <- c(A = "ACTGGCCA", B = "ACTGGCCT", C = "ACTGTCCA" )
cat(fly_seq, sep ="\n")
```

**Make a single string that looks like "x1 + x2 + x3 + x4 + x5" using `paste` or for stringi it is `stri_join()` and `str_c` in `stringr` **

```{r}
paste("x", 1:5, sep = "", collapse = " + ")
```


**I have created an object of DNA sequences called "seqs" how long is each sequence? How many sequences are there? `nchar()` in base, and `str_length()`**

```{r, echo = F}
seq_gen <- function() {
   x <- sample(c("A", "C", "T", "G"), 
           size = rbinom(100, 100, 0.8),
           replace =T)
   paste0(x, collapse="")    
}
seqs <- replicate(20, seq_gen())
```

```{r}
nchar(seqs)
length(seqs)
```

**Words in the origin of species**

```{r, message = F}
library(gutenbergr)
library(tidyverse)

books <- gutenberg_works(grepl(title, 
                               pattern = "On the origin of species",
                         ignore.case = TRUE))

species <- gutenberg_download(1228)
str(species)
head(species$text)
species_text <- species$text
```

**How many lines are empty, check using the `stri_isempty`. What proportion of lines is this?**

```{r}
empty <- stri_isempty(species_text)
mean(empty)
```

**How often does the word introduction appear? Use `grep` or `str_subset`**

```{r}
grep(species_text, pattern = "introduction", ignore.case= T)
```

**How would you check the lines?**

```{r}
for (x in grep(species_text, pattern = "introduction", ignore.case= T)) {
  print(species_text[x])
}
```

Let's use slice in dplyr to remove introduction and INDEX as follows (or really just keeping everything in between)


```{r}
species_slice <- species %>%
  slice(grep("^INTRODUCTION\\.", text):(grep("^INDEX\\.", text))-1)
```

**Identifying chapters**

Each chapter begins like this "1. VARIATION UNDER DOMESTICATION". Can you identify a regular expression that matches this?

```{r}
grep("^[0-9]+\\.", species_slice$text, value=TRUE)
```


**removing blank lines etc...**

Use nzchar (which returns FALSE is the string is empty) to find empty lines

```{r}
length(species_slice$text)
sum(nzchar(species_slice$text))
```

**We can filter on those lines. Fill in the missing parts.**

```{r}
species_filtered <- species %>%
  slice(grep("^INTRODUCTION\\.", text):(grep("^INDEX\\.", text))-1) %>%
  filter(nzchar(text)) %>%
  mutate(linenumber = row_number(),
         chapter = cumsum(grepl("^[0-9]+\\.", text)))
```

**`tidytext`**
We could continue to make all of the changes by ourselves, but let's use tidytext to help us

```{r}
species_tidy <- species_filtered %>%
    unnest_tokens(word, text)
```

Now we remove "stop words"

```{r}
species_tidy <- species_tidy %>%
  anti_join(stop_words)
```

**Find the most common words in the book?**

```{r}
species_count <- species_tidy %>% count(word) %>% arrange(desc(n))
species_count
```

**How about a word cloud of the top 50 words**

```{r, message = F}
library(wordcloud)
set.seed(57)
ran_col <- sample(colors(), size = 300)
wordcloud(species_count$word, species_count$n, max.words=300, min.freq=30 ,random.color=T, colors = ran_col, scale=c(5,.3), random.order = F)
```


