---
title: "Lesson 2 In Class Exercise Part 1"
author: "Lucy Zhang"
date: "October 29, 2018"
output: 
  html_document: 
    keep_md: yes
---

```{r}
read_1 <- "CGCGCAGTAGGGCACATGCCAGGTGTCCGCCACTTGGTGGGCACACAGCCGATGACGAACGGGCTCCTTGACTATAATCTGACCCGTTTGCGTTTGGGTGACCAGGGAGAACTGGTGCTCCTGC"
read_2 <- "AAAAAGCCAACCGAGAAATCCGCCAAGCCTGGCGACAAGAAGCCAGAGCAGAAGAAGACTGCTGCGGCTCCCGCTGCCGGCAAGAAGGAGGCTGCTCCCTCGGCTGCCAAGCCAGCTGCCGCTG"
read_3  <- "CAGCACGGACTGGGGCTTCTTGCCGGCGAGGACCTTCTTCTTGGCATCCTTGCTCTTGGCCTTGGCGGCCGCGGTCGTCTTTACGGCCGCGGGCTTCTTGGCAGCAGCACCGGCGGTCGCTGGC"
```
**Question 1. what species are these sequences from?**\
*D. melanogaster*

_____

**Question 2. Put all three of these reads into a single object (a vector).  What class will the vector `reads` be? Check to make sure! How many characters are in each read (and why does `length()` not give you what you want.. try...)**\
```{r}
reads <- c(read_1, read_2, read_3)
class(reads)
length(reads)
```
Reads only returns the number of items in the vector not the length of the strings in each item of the vector.

_____

**Question 3. Say we wanted to print each character (not the full string) from read_1, how do we do this using a for loop? You may wish to look at a function like `strsplit()` to accomplish this (there are other ways.)**\
Replace the blanks.
```{r}
read_1_split <- strsplit(read_1, split = "", fixed = T) 
print(read_1_split)
```

_____
**Question 4. What kind of object does this return? How might we make it a character vector again?**\
```{r}
class(read_1_split)
```
This returns a `list` object.
```{r}
read_1_vec <- unlist(read_1_split)
class(read_1_vec)
```
_____
**Question 5. How about if we wanted the number of occurrences of each base? Or better yet, their frequencies? You could write a loop, but I suggest looking at the help for the `table()` function... Also keep in mind that for for most objects `length()` tells you how many elements there are in a vector. For lists use `lengths()` (so you can either do this on a character vector or a list, your choice)**\
```{r}
print(table(read_1_split)/lengths(read_1_split), digits = 2)
```
_____
**Question 6. How would you make this into a nice looking function that can work on either  a list or vectors of characters? (Still just for a single read)**\
```{r}
count <- function(x) {
  if (class(x) == "character") {
    table <- table(x)/length(x)
  } else if (class(x) == "list") {
    unlist(x)
    table <- table(x)/lengths(x)
  } else {
    print("Invalid data type")
  }
  return(table)
}
print(count(read_1_split), digits = 2)
print(count(read_1_vec), digits = 2)
```
_____
**Question 7. Now how can you modify your approach to do it for an arbitrary numbers of reads? You could use a loop or use one of the apply like functions (which one)?**\
```{r}
count_2 <- function(x) {
  if (class(x) == "character" & length(x) == 1) {
    x <- strsplit(x, split = "", fixed = T) 
    x <- unlist(x)
    table <- table(x)/length(x)
  } else if (class(x) == "character" & length(x) != 1) {
    table <- table(x)/length(x)
    } else if (class(x) == "list") {
    unlist(x)
    table <- table(x)/length(x)
  } else {
    print("Invalid data type")
  }
  return(table)
}
print(sapply(reads, count_2, USE.NAMES = F), digits = 2)
```
_____
**Question 8. Can you revise your function so that it can handle the input of either a string as a single multicharacter vector, *or* a vector of individual characters *or* a list? Try it out with the vector of three sequence reads (`reads`).**\
```{r}
count_3 <- function(x) {
  if (class(x) == "character" & length(x) == 1) {
    x <- strsplit(x, split = "", fixed = T) 
    x <- unlist(x)
    table <- table(x)/length(x)
  } else if (class(x) == "character" & length(x) != 1) {
      for (item in x) {
        if (length(item) == 1) {
          split <- strsplit(x, split = "", fixed = T)
          table <- table(split)/length(split)
        } else {
           table <- table(x)/length(x)
        }
      }
    } else if (class(x) == "list") {
    unlist(x)
    table <- table(x)/length(x)
  } else {
    print("Invalid data type")
  }
  return(table)
}
print(count_3(reads), digits = 2)
```