---
title: "Lesson 2 In Class Exercise Part 2"
author: "Lucy Zhang"
date: "October 29, 2018"
output: 
  html_document: 
    keep_md: yes
    number_sections: yes
    toc: yes
---
**1. What happens to `x` for each of these functions after you run them? Please explain what is happening in each case. i.e. Why do these behave differently?**\

In the first case, x is still found in the namespace because the for-loop is executed in the global environment, in the second case, x is not found because it is contained within the function.

_____

**2. Write a for loop that generates a vector which is the square of all integers from 1 to 10000. Make this all happen within the loop. i.e. We are going to use the `system.time()` function to monitor how long it takes. This is pretty coarse, and a way better way of doing this is with the `microbenchmark` package (which is worth learning), but for now this function will do.**\
```{r}
rm(x)
x <- c()
system.time(for (num in 1:10000) x[num] <- num^2 )
```
_____

**3.  Now do the same thing as in question 2, but *pre-allocate* memory for the vector `x` (or whatever you want to call it) to store the 10000 values *before* running the `for` loop. You don't need to do anything differently, just initialize a vector of length 10000. Which runs faster? Please explain why?**\
```{r}
rm(x)
x <- c()
mySeq <- c(1:10000)
system.time(for (num in mySeq) x[num] <- num^2 )
```
The first case runs faste because R stores the values in virtual memory and is faster to retrieve than physical memory, which is the second case.

_____

**4. What is a very "`R`" to make this (question 2 and 3) even faster? (i.e. think about vectorization)**\
```{r}
rm(x)
x <- 1:10000
system.time(x <- x^2)
```
